---
title: "Lab 11 - Smoking during pregnancy"
author: "Ryan Wheat"
date: "05/08/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

### Exercise 1

Each case is one baby in North Carolina. Thus, each variable describes something that applies to that baby: the age of their mother/father, their weight, etc. There are a total of 1,000 babies in this dataframe.


```{r exercise-1}

data(ncbirths)
view(ncbirths)

```

### Exercise 2

The average white baby weighs 7.25 pounds.

```{r exercise-2}

ncbirths_white <- ncbirths %>%
  filter(whitemom == "white")

mean_weight <- mean(ncbirths_white$weight)

mean_weight
```

### Exercise 3

We need to simulate the population mean because we only have a sample, and not the population. 

### Exercise 4

These results reveal that white babies in NC now weigh less (M = 7.25) than they did in 1995 (M = 7.43), p = .001.

```{r exercise-4}

#initial bootstrap

set.seed(1000)

birth_bootstrap <- ncbirths_white %>%
  specify(response = weight) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")

#create null distribution

null_center <- 7.43 - mean(birth_bootstrap$stat, na.rm = TRUE)

null_dist <- birth_bootstrap %>%
  mutate(stat = stat + null_center)

#visualize

ggplot()+
  geom_histogram(data = null_dist, aes(x = stat), fill = "skyblue", color = "black", alpha = 0.25)+
  geom_histogram(data = birth_bootstrap, aes(x = stat), fill = "navyblue", color = "black", alpha = 0.25)+
  theme_bw()

#test

observed_weight <- ncbirths_white %>%
  specify(response = weight) %>%
  calculate(stat = "mean")

get_p_value(null_dist, obs_stat = observed_weight, direction = "two-sided")
```

### Exercise 5

This boxplot shows that the babies from smokers vs. nonsmokers in the dataframe generally weighed similarly, but that there are more outliers in the nonsmoker group.

```{r boxplot-visual}

ncbirths %>%
  filter(!is.na(habit) & !is.na(weight)) %>%
  ggplot(aes(x = habit, y = weight))+
  geom_boxplot() +
  theme_bw()

```

### Exercise 6

```{r cleaned-data}

ncbirths_cleaned <- ncbirths %>%
  filter(!is.na(habit) & !is.na(weight))

```

### Exercise 7

```{r mean-diff}

ncbirths_cleaned %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight))
```
### Exercise 8

H0: uSmoking = uNonSmoking
H1: uSmoking =/= uNonSmoking

### Exercise 9

It seems that smokers had smaller babies than non-smokers (p = .002).

```{r inital-bootstrap}

set.seed(1000)

smoker_bootstrap <- ncbirths_cleaned %>%
  specify(weight ~ habit) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "diff in means", order = c("smoker", "nonsmoker"))
```

```{r null-and-plot}

#set null distribution

null_center_new <- 0 - mean(smoker_bootstrap$stat, na.rm = TRUE)

null_smoke <- smoker_bootstrap %>%
  mutate(stat = stat + null_center_new)

ggplot()+
  geom_histogram(data = null_smoke, aes(x = stat), fill = "skyblue", color = "black", alpha = 0.2)+
  geom_histogram(data =smoker_bootstrap, aes(x = stat), fill = "navyblue", color = "black", alpha = 0.2)+
  theme_bw() +
  labs(x = "Mean Difference",
       y = "Count")
```

```{r inferential-test}

weight_diff <- ncbirths_cleaned %>%
  specify(weight ~ habit) %>%
  calculate(stat = "diff in means", order = c("smoker", "nonsmoker"))

get_p_value(null_smoke, obs_stat = weight_diff, direction = "both")
```

### Exercise 10

I am going to seperate mature from young mothers by doing a median split. 

```{r median-split}

ncbirths_cleaned <- ncbirths_cleaned %>%
  mutate(mother_agegroup = case_when(mage > median(mage) ~ "Mature Mother",
         mage <= median(mage) ~ "Younger Mother"))

```

### Exercise 11

According to our results, there does not seem to be a difference in birth weight between older and younger mothers.

```{r test}

set.seed(1000)

weight_age <- ncbirths_cleaned %>%
  specify(response = lowbirthweight, explanatory = mother_agegroup, success = "not low") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "diff in props", order = c("Younger Mother", "Mature Mother"))

#set null

null_center_three <- 0 - mean(weight_age$stat, na.rm = TRUE)

weight_age_null <- weight_age %>%
  mutate(stat = stat - null_center_three)

ggplot()+
  geom_histogram(data = weight_age_null, aes(x = stat), fill = "skyblue", color = "black", alpha = 0.5)+
  geom_histogram(data = weight_age, aes(x = stat), fill = "navyblue", color = "black", alpha = 0.5)+
  theme_classic()

#test

differences <- ncbirths_cleaned %>%
  specify(response = lowbirthweight, explanatory = mother_agegroup, success = "not low") %>%
  calculate(stat = "diff in props", order = c("Younger Mother", "Mature Mother"))
            
get_p_value(weight_age_null, obs_stat = differences, direction = "both")

```

### Exercise 13

If we had data from all mothers, we would expect the mean weight of babies born to mature mothers to be anywhere from 3% lighter to 4% heavier than those born to younger mothers.

```{r last-thing-in-r-class}

get_ci(weight_age, level = 0.95)

```

